\subsection{Sourcing}
The primary dataset was acquired from a publicly available open-source Kaggle dataset comprising multilingual songs from various genres. Specifically, the dataset was compiled from aggregated external sources such as the Spotify API and Kaggle datasets: \textit{150K Lyrics Labelled with Spotify Valence}, \textit{Lyrics Musics}, and \textit{AZLyrics Song Lyrics}. The final dataset was downloaded and processed directly from \textit{Kaggle}. However, since the dataset included additional attributes and records unrelated to the task, further preprocessing was necessary. 

\subsection{Preprocessing}
Initial preprocessing involved removing non-English songs, additional genres and invalid records from the dataset. Within Kaggle, the dataset is separated into two CSV files: test.csv and train.csv. To acquire the maximum  number of data points, the data from each csv file (train and test) were downloaded and merged into one csv file. Afterward. Non-English songs were filtered and removed using the "language" attribute column of the dataset. The initial dataset contained five default attributes: lyrics, genre, language, artist name and song title. Since some of these attributes are not essential for annotation and modelling, the dataset was filtered to only retain 2 attributes: genre and lyrics. During this process, additional genres (other than Pop and Metal) were removed. Moreover, any records with a or empty "lyrics" attribute were removed to maintain completeness. Unique IDs (integer values, based on order) for each record (music genre + lyric) were added as an attribute to identify each record's lyrics and ground-truth label (genre). The final preprocessed dataset consisted of three attributes:  lyric\_id, genre and lyrics. 

\subsection{Size and Distribution}
The preprocessed dataset contained 105,443 records (lyric\_id, genre, lyrics). However, due to the initial Kaggle dataset being compiled from external sources (i.e. Spotify API), there was significant class imbalance between the two classes: Pop (86298) and Metal (19133). To address this class imbalance, stratification and downsampling were utilized.

\subsubsection{Downsampling}
Primarily, downsampling was conducted to ensure the quantity of samples of each class is equal. Specifically, the process reduced the amount of samples of the Pop class from 86298 to 19133, which is represents the maximum amount of samples in the Metal class. This was done through random down sampling. First, the process finds the minority class size in the dataset (which is Metal with only 19133 samples). Then, random sampling is conducted to gather the samples for each class. This results in both classes (Pop and Metal) containing 19133 samples. Although a generic threshold could have been used, downsampling to the maximum available threshold (19133) was done to maximize data for downstream model training. Most importantly, this process was applied to remove the majority class bias (the majority of the Pop samples) in the dataset and create a more balanced downstream data distribution.

\subsubsection{Stratification}
Stratification was applied to preserve the ratio of the two classes in each dataset split. Firstly, the records of the balanced dataset (following downsampling) were randomly shuffled to prevent ordering bias and provide a random distribution before splitting. Afterward, during the train/test/validation split, stratification was applied to the dataset to ensure equal class ratios (pop vs. metal) on all splits. This resulted in the removal of class imbalance during model training and evaluation.

\subsubsection{Annotation}
Since the initial dataset consisted of ground-truth labels (genre attribute), additional annotation was not performed. As a result, for downstream training and evaluation, the "genre" attribute was utilized as the defined label for each record (song lyrics). 

