The best performing model was the Logistic Regression model, compared to the pretrained models tested. When analyzing the model's failures, a thematic challenge emerges regarding context and metaphor. The model was built using a robust feature pipeline that combined word and character n-gram TF-IDF vectors with custom numeric features (like type-token ratio, line counts, and punctuation usage). However, in instances such as song ID 114 and ID 1123, the model incorrectly guessed Pop for Metal songs because the lyrics centered on vulnerability and romance. Lines like "Light me a candle, I'm coming home / Then leave it on your side where I belong" (ID 114) and "Heartbreak and romance / It's better to be numb if you're cracking at the heart" (ID 1123) were misclassified because terms like "home," "heartbreak" and "romance" are statistically dominated by Pop music in the TF-IDF training data, overriding structural features and masking the heavier instrumentation that actually When the model incorrectly guessed Metal for a Pop song, as seen in ID 83472 and ID 60071, the lyrics relied heavily on aggressive metaphors. Lines like "Take away my prime from the mouth of the lion / Savage jaws are the price we pay" (ID 83472) and "Every spark has an ending like a burning fuse" (ID 60071) tricked the model because words like "lion," "savage," "jaws," and "burning" strongly trigger Metal classifications. The primary challenge here is that despite including structural features, the dominant TF-IDF vectors take words literally, failing to recognize when an aggressive word is just a metaphor for a pop breakup or when a soft word is sung over heavy, distorted guitars.

Despite these edge cases, the Logistic Regression pipeline is effective when songs lean into standard genre tropes. For example, a successful Pop classification for song ID 99344 ("head buster off top / grab the chopper and chop / at the top is my spot") demonstrates the model capturing rhythmic, upbeat language and modern slang typical of mainstream hip-hop/pop crossovers. On the other hand, a successful Metal classification for song ID 67457 ("Remnants strewn upon the ground / Drag the carcass withered and brown / Sockets crushed powder remaining") shows the model correctly identifying the overtly dark, morbid, and gothic vocabulary that serves as a staple of the metal genre. In these cases, the semantic meaning perfectly aligns with the literal definitions of the words, allowing the TF-IDF features to confidently and accurately predict the genre based purely on vocabulary frequency.

To improve the robustness of this model and eliminate the literal interpretation blindspots, future iterations should focus on moving beyond strictly frequency-based representations. Additionally, because lyrics only tell half the story, utilizing external APIs to include acoustic features such as energy, tempo, and instrumental features, alongside the existing custom numerical features would immediately resolve ambiguities between songs with similar lyrical themes but vastly different musical styles.